Parsing Postfix log files

### 2009-02-23 doc/logparser-acronyms.tex
Added a graph that utilises the data about the number of rules tried, lines     parsed, etc. Added examples in background section to deal with remaining XXX items. Dealt with some other XXX items, and added some more. Improved the Makefile and fixed some dependency mistakes. Started adding "last checked" dates to all URLs. Improved the introduction in the architecture chapter.

\newglossaryentry{DNSBL}{
    type=\acronymtype,
    name={DNSBL},
    description={DNS Blacklist},
    first={DNS Blacklist (DNSBL)}
}
\newglossaryentry{DNSBL glossary}{
    name={DNS Blacklist},
    description={
        A DNS Blacklist~\cite{Wikipedia-DNSBL} is a simple collaborative
        anti-spam technique used to reject or penalise email sent from mail
        servers reported to have sent large volumes of spam.  To use a
        DNSBL the mail server makes a DNS request incorporating the IP
        address of the client; if the requested hostname exists the client
        is on the DNSBL, and the mail server can decide what course of
        action to take.
    }
}

@@

\newglossaryentry{SQL}{
    type=\acronymtype,
    name={SQL},
    description={Structured Query Language},
    first={Structured Query Language (SQL)}
}
\newglossaryentry{SQL glossary}{
    name={Structured Query Language},
    description={
        The standard language used for database querying, modification, and
        maintenance.  Some information and history can be found at
        \url{http://en.wikipedia.org/wiki/SQL}; a good introduction can be
        found at \url{http://philip.greenspun.com/sql/}.
    }
}

>>>>

\newglossaryentry{DNSBL}{
    type=\acronymtype,
    name={DNSBL},
    description={DNS Blacklist},
    first={DNS Blacklist (DNSBL)}
}
\newglossaryentry{DNSBL glossary}{
    name={DNS Blacklist},
    description={
        A DNS Blacklist~\cite{Wikipedia-DNSBL} is a simple collaborative
        anti-spam technique used to reject or penalise mail sent from mail
        servers believed to be the source of large volumes of spam.  The
        criteria used when deciding whether or not to include an IP address
        on a DNSBL vary widely between DNSBLs, so before using one it is
        very important to check their listing policies.  To use a DNSBL the
        mail server makes a DNS request incorporating the IP address of the
        client; if the requested hostname exists the client is on the
        DNSBL, and the mail server can decide what course of action to
        take.
    }
}

@@

\newglossaryentry{SQL}{
    type=\acronymtype,
    name={SQL},
    description={Structured Query Language},
    first={Structured Query Language (SQL)}
}
\newglossaryentry{SQL glossary}{
    name={Structured Query Language},
    description={
        The standard language used for database querying, modification, and
        maintenance.  Some information and history can be found at
        \urlLastChecked{http://en.wikipedia.org/wiki/SQL}{XXX}; a good
        introduction can be found at
        \urlLastChecked{http://philip.greenspun.com/sql/}{XXX}.
    }
}

### 2009-02-23 doc/logparser-architecture.tex
Added a graph that utilises the data about the number of rules tried, lines     parsed, etc. Added examples in background section to deal with remaining XXX items. Dealt with some other XXX items, and added some more. Improved the Makefile and fixed some dependency mistakes. Started adding "last checked" dates to all URLs. Improved the introduction in the architecture chapter.

To avoid cluttering the explanation of the parser architecture with details
of the difficulties encountered while parsing Postfix log files, the
description of the parser architecture has been separated from the
description of \parsername{}, the implementation developed for this
project.  This chapter presents the architecture that has been designed and
developed for this project, beginning with the overall architecture and
design, followed by the three components of the architecture: Framework,
Actions, and finally Rules.  The discussion in this chapter centers on the
theoretical, XXX EXTEND WHEN THE CHAPTER IS FINISHED\@.
The practical difficulties of writing a parser for Postfix log files are
covered in detail in \sectionref{Postfix Parser Implementation}.

@@

\subsection{Rule Characteristics}

\label{rule characteristics in architecture}

@@

\subsection{Architecture Characteristics}

>>>>

To avoid cluttering the explanation of the parser architecture with the
details involved in parsing Postfix log files and the description of the
parser implemented for this project, the two have been separated.  This
chapter presents the architecture designed and developed for this project,
beginning with the overall architecture and design, followed by the three
components of the architecture: Framework, Actions, and Rules, and
finishes by describing the characteristics of this architecture.

This chapter centers on the theoretical, implementation-independent aspects
of the architecture; the practical difficulties of writing a parser for
Postfix log files are covered in detail in \sectionref{Postfix Parser
Implementation}.

@@

@@

\section{Architecture Characteristics}

### 2009-02-23 doc/logparser-background.tex
Added a graph that utilises the data about the number of rules tried, lines     parsed, etc. Added examples in background section to deal with remaining XXX items. Dealt with some other XXX items, and added some more. Improved the Makefile and fixed some dependency mistakes. Started adding "last checked" dates to all URLs. Improved the introduction in the architecture chapter.

Postfix is a \glsfirst{MTA} with the following design aims
(in order of importance): security, flexibility of configuration,
scalability, and high performance.  It features extensive optional
anti-spam restrictions, allowing an administrator to employ those
restrictions which they judge suitable for their site's needs, rather than
a fixed set chosen by Postfix's author.  These restrictions can be
selectively applied, combined, and bypassed on a per-client, per-recipient,
or per-sender basis, allowing different levels of stricture and
permissiveness.  Postfix leverages simple lookup tables to support
arbitrarily complicated user-defined sequences of restrictions and
exceptions.  Administrators can supply their own rejection messages to make
it clear to senders exactly why their mail was rejected.  Policy servers
(\sectionref{policy servers}) provide a simple way to write new
restrictions without having to modify Postfix's source code.  Unfortunately
this flexibility has a cost: complexity in the log files generated.
Although it is relatively simple in most cases to use standard Unix text
processing utilities to determine the fate of an individual email,
following the journey an email takes through Postfix can be quite
difficult.  For most mails the journey is simple and brief, but the
remaining minority can have quite complex journeys (see
\sectionref{complications} for details).

Postfix's design follows the Unix philosophy of \textit{``Write programs
that do one thing and do it well''\/}~\cite{unix-philosophy}, and it is
separated into multiple programs that each perform one of the tasks
required of an \acronym{MTA}\@: receive mail, send mail, local delivery of
mail, etc.\ (full details can be found in~\cite{postfix-overview}).  Each
log line contains the name of the Postfix component that produced it, and
this information is used when determining which rules should be used to
parse each log line (see \sectionref{rule characteristics} for details).
Postfix's design is strongly influenced by security concerns: those
components that interact with other hosts are not
privileged,\footnote{Privilege means the power to perform actions that are
limited to the administrator, and not available to ordinary users.} so bugs
in those components will not give an attacker extra privileges; those
components that are privileged do not interact with other hosts, making it
much more difficult for an attacker to exploit any bugs that may exist in
those components.

@@

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno}, or the name of another restriction to be
evaluated; other less commonly used results are possible as described
in~\cite{smtpd_access_readme,smtpd_per_user_control,policy-servers}. The
meaning of \textit{permit\/} and \textit{reject\/} is fairly obvious;
\textit{dunno\/} means to stop evaluating the current restriction and
continue processing the remainder of the restriction list, allowing
exceptions to more general rules.  XXX GIVE AN EXAMPLE\@.  When the result
is the name of another restriction Postfix will evaluate that restriction,
allowing restrictions to be chosen based on the client \acronym{IP}
address, client hostname, HELO hostname, sender address, or recipient
address.  E.g.\ the administrator may require that all clients on the local
network have valid DNS entries, to prevent people sending mail from unknown
machines XXX EXPAND THIS WITH DETAILS\@.

The administrator can define new restrictions as a list of existing
restrictions, allowing arbitrarily long and complex sequences of lookups,
restrictions and exceptions.  Postfix tries to protect the administrator
from misconfiguration in as far as is reasonable, e.g.\ the restriction
\texttt{check\_helo\_mx\_access} cannot cause a mail to be accepted,
because the parameter it checks (the hostname given in the HELO command) is
under the control of the remote client.  Despite this, it is possible for
the administrator to make catastrophic mistakes, e.g.\ rejecting all mail
--- the administrator must be cognisant of the effects their configuration
changes will have.  This is similar to one of UNIX's design philosophies:
\textit{``UNIX was not designed to stop its users from doing stupid things,
as that would also stop them from doing clever
things''\/}~\cite{unix-philosophy}.

>>>>

Postfix is a \glsfirst{MTA} with the following design aims (in order of
importance): security, flexibility of configuration, scalability, and high
performance.  It features extensive optional anti-spam restrictions,
allowing an administrator to employ those restrictions which they judge
suitable for their site's needs, rather than a fixed set chosen by
Postfix's author.  These restrictions can be selectively applied, combined,
and bypassed on a per-client, per-recipient, or per-sender basis, allowing
different levels of stricture and permissiveness.  Postfix leverages simple
lookup tables to support arbitrarily complicated user-defined sequences of
restrictions and exceptions.  Administrators can supply their own rejection
messages to make it clear to senders exactly why their mail was rejected.
Policy servers (\sectionref{policy servers}) provide a simple way to write
new restrictions without having to modify Postfix's source code.
Unfortunately this flexibility has a cost: complexity in the log files
generated.  Although it is relatively simple in most cases to use standard
Unix text processing utilities to determine the fate of an individual mail,
following the journey a mail takes through Postfix can in some cases be
quite difficult.  For most mails the journey is simple and brief, but the
remainder can have quite complex journeys (see \sectionref{complications}
for details).

Postfix's design follows the Unix philosophy of \textit{``Write programs
that do one thing and do it well''\/}~\cite{unix-philosophy}, and it is
separated into multiple programs that each perform one of the tasks
required of an \acronym{MTA}\@: receive mail, send mail, local delivery of
mail, etc.\ (full details can be found in~\cite{postfix-overview}).  Each
log line contains the name of the Postfix component that produced it, and
this information is used when determining which rules should be used to
parse each log line (see \sectionref{rules in architecture} for details).
Postfix's design is strongly influenced by security concerns: those
components that interact with other hosts are not
privileged,\footnote{Privilege means the power to perform actions that are
limited to the administrator, and not available to ordinary users.} so bugs
in those components will not give an attacker extra privileges; those
components that are privileged do not interact with other hosts, making it
much more difficult for an attacker to exploit any bugs that may exist in
those components.

@@

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno}, or the name of another restriction to be
evaluated; other less commonly used results are possible as described
in~\cite{smtpd_access_readme,smtpd_per_user_control,policy-servers}. The
meaning of \textit{permit\/} and \textit{reject\/} is fairly obvious;
\textit{dunno\/} means to stop evaluating the current restriction and
continue processing the remainder of the restriction list, allowing
exceptions to more general rules.  E.g.\ the general rule is that machines
on the local network must authenticate before sending mail, except for the
web server, because the legacy applications running on it lack
authentication support; this configuration example shows how:

\begin{verbatim}
  main.cf:
  smtpd_client_restrictions =
    ...
    permit_sasl_authenticated,
    check_client_access /etc/postfix/allow_webserver.cidr,
    ...

  /etc/postfix/allow_webserver.cidr:
    192.0.2.80/32   dunno
    192.0.2.0/24    reject "Please authenticate to send mail"
\end{verbatim}

That example also shows how easy it is to supply a custom rejection
message.  When the result is the name of another restriction Postfix will
evaluate that restriction, allowing restrictions to be chosen based on the
client \acronym{IP} address, client hostname, HELO hostname, sender
address, or recipient address.  E.g.\ the administrator may require that
all clients on the local network have valid DNS entries, to prevent people
sending mail from unknown machines; one example of how to achieve this is:

\begin{verbatim}
  main.cf:
  smtpd_client_restrictions =
    ...
    check_client_access /etc/postfix/require_dns_entries.cidr,
    ...

  /etc/postfix/require_dns_entries.cidr:
    192.0.2.0/24    reject_unknown_client_hostname
\end{verbatim}

The administrator can define new restrictions as a list of existing
restrictions, allowing arbitrarily long and complex sequences of lookups,
restrictions, and exceptions.  Postfix tries to protect the administrator
from misconfiguration in as far as is reasonable, e.g.\ the restriction
\texttt{check\_helo\_mx\_access} cannot cause a mail to be accepted,
because the parameter it checks (the hostname given in the HELO command) is
under the control of the remote client.  Despite this, it is possible for
the administrator to make catastrophic mistakes, e.g.\ rejecting all mail
--- the administrator must be cognisant of the effects their configuration
changes will have.  This is similar to one of UNIX's design philosophies:
\textit{``UNIX was not designed to stop its users from doing stupid things,
as that would also stop them from doing clever
things''\/}~\cite{unix-philosophy}.

### 2009-02-23 doc/logparser-implementation.tex
Added a graph that utilises the data about the number of rules tried, lines     parsed, etc. Added examples in background section to deal with remaining XXX items. Dealt with some other XXX items, and added some more. Improved the Makefile and fixed some dependency mistakes. Started adding "last checked" dates to all URLs. Improved the introduction in the architecture chapter.

    \item [sender] The sender's email address.  Because multiple mails may
        be delivered during one connection, there may be different sender
        addresses in the results for one connection; however there should
        not be different sender addresses in the results for one email.
        Mails sent over the same connection can be distinguished by their
        queueid.

>>>>

    \item [sender] The sender's email address.  Because multiple mails may
        be delivered during one connection, there may be different sender
        addresses in the results for one connection; however there should
        not be different sender addresses in the results for one mail.
        Mails sent over the same connection can be distinguished by their
        queueid.

### 2009-02-23 doc/logparser-introduction.tex
Added a graph that utilises the data about the number of rules tried, lines     parsed, etc. Added examples in background section to deal with remaining XXX items. Dealt with some other XXX items, and added some more. Improved the Makefile and fixed some dependency mistakes. Started adding "last checked" dates to all URLs. Improved the introduction in the architecture chapter.

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam measures are working.  The more
adventurous will have generated statistics to show how successful each of
their anti-spam measures has been in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this kind of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\acronym{DNSBL}\footnote{This document is supplied with a glossary, see
\textsection\ref{Glossary}.} returns a different explanatory message;
policy servers may log different messages depending on the characteristics
of the connection.  There are many ways in which the log lines may differ
between servers, even within the same organisation: servers may be
configured differently, or running different versions of Postfix.  This
paper documents the difficult process of parsing Postfix log files,
presenting \parsername{}, a program that parses Postfix log files and
places the resulting data into a database for later use.  The gathered data
can then be used to optimise current anti-spam measures, provide a baseline
to test new anti-spam measures against, or to produce statistics showing
how effective those measures are.  Numerous other uses are possible for
such data: improving server performance by identifying troublesome
destinations and reconfiguring appropriately; identifying regular high
volume uses (e.g.\ customer newsletters) and restricting those uses to
off-peak times; detecting virus outbreaks that propagate via email; as a
base for billing customers on a shared server.  Preserving the raw data
enables users to develop a multitude of uses far beyond those conceived of
by the author.

>>>>

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam measures are working.  The more
adventurous will have generated statistics to show how successful each of
their anti-spam measures has been in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this kind of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\acronym{DNSBL}\footnote{This document is supplied with a glossary, see
\textsection\ref{Glossary}.} returns a different explanatory message;
policy servers may log different messages depending on the characteristics
of the connection.  There are many ways in which the log lines may differ
between servers, even within the same organisation: servers may be
configured differently, or running different versions of Postfix.  This
paper documents the difficult process of parsing Postfix log files,
presenting \parsername{}, a program that parses Postfix log files and
places the resulting data into a database for later use.  The gathered data
can then be used to optimise current anti-spam measures, provide a baseline
to test new anti-spam measures against, or to produce statistics showing
how effective those measures are.  Numerous other uses are possible for
such data: improving server performance by identifying troublesome
destinations and reconfiguring appropriately; identifying regular high
volume uses (e.g.\ customer newsletters) and restricting those uses to
off-peak times; detecting virus outbreaks that propagate via mail; as a
base for billing customers on a shared server.  Preserving the raw data
enables users to develop a multitude of uses far beyond those conceived of
by the author.

### 2009-02-23 doc/logparser-other-parsers-review.tex
Fixed some 'overfull hbox' problems.
Added a graph that utilises the data about the number of rules tried, lines     parsed, etc. Added examples in background section to deal with remaining XXX items. Dealt with some other XXX items, and added some more. Improved the Makefile and fixed some dependency mistakes. Started adding "last checked" dates to all URLs. Improved the introduction in the architecture chapter.

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface provides search functionality, generating statistics and
graphs in real time, a facility not provided by \parsername{}.  Splunk
allows quite complicated searches, based on the fields extracted by Splunk
(described later) or the full text of the log line, though it is not
possible to search on partial words.  Search results are not available for
use with other tools.  Searches can be saved for reuse; saved searches can
be run periodically and the results emailed to a recipient or sent to an
external program for further processing.  The author was unable to save
searches, though that may have been because of limitations in the free
version.  The data store is not available for use by external programs,
whereas \parsername{} provides the database and leaves it to the user to
utilise it without limit or restriction.  The web interface is optimised
for interactive use rather than automated queries, and it does not appear
to be possible to write independent tools to utilise the Splunk database.
Some additional Postfix reports are supposedly available on
\url{http://www.splunkbase.com/}, but the author was unable to find any
Postfix reports, or indeed reports for any other log file types: every
category was empty, even those that the web site claimed had numerous
reports available.  Many types of graphs can be generated, though most are
variations of a bar or pie chart, except the bubble and heatmap graphs.  It
is easy to drill down through the graphs to select a portion of the data
(e.g.\ select the hour with the largest number of events, then select a
particular host, and finally a specific sender address).  All searches
performed using the indexed data returned reasonable results.

@@

\parserblurb{%
    As any good system administrator knows, there's a lot more to keep
    track of in an active network than just webservers. Lire is hands down
    the most versatile log analysis software available today. Lire not only
    keeps you informed about your HTTP, FTP, and mail traffic, it also
    reports on your firewalls, your print servers, and your DNS activity.
    The ever growing list of Lire-supported services clearly outstrips any
    other software, in large part thanks to the numerous volunteers who
    have pioneered many new services and features. Lire is a total solution
    for your log analysis needs.
}
{http://logreport.org/lire.html}
{2009/01/11}

@@

    \item The pair of correspondents exchanged the highest number of
        emails.

>>>>

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface provides search functionality, generating statistics and
graphs in real time, a facility not provided by \parsername{}.  Splunk
allows quite complicated searches, based on the fields extracted by Splunk
(described later) or the full text of the log line, though it is not
possible to search on partial words.  Search results are not available for
use with other tools.  Searches can be saved for reuse; saved searches can
be run periodically and the results mailed to a recipient or sent to an
external program for further processing.  The author was unable to save
searches, though that may have been because of limitations in the free
version.  The data store is not available for use by external programs,
whereas \parsername{} provides the database and leaves it to the user to
utilise it without limit or restriction.  The web interface is optimised
for interactive use rather than automated queries, and it does not appear
to be possible to write independent tools to utilise the Splunk database.
Some additional Postfix reports are supposedly available on
\url{http://www.splunkbase.com/}, but the author was unable to find any
Postfix reports, or indeed reports for any other log file types: every
category was empty, even those that the web site claimed had numerous
reports available.  Many types of graphs can be generated, though most are
variations of a bar or pie chart, except the bubble and heatmap graphs.  It
is easy to drill down through the graphs to select a portion of the data
(e.g.\ select the hour with the largest number of events, then select a
particular host, and finally a specific sender address).  All searches
performed using the indexed data returned reasonable results.

@@

\sloppy{}%
\parserblurb{%
    As any good system administrator knows, there's a lot more to keep
    track of in an active network than just webservers. Lire is hands down
    the most versatile log analysis software available today. Lire not only
    keeps you informed about your HTTP, FTP, and mail traffic, it also
    reports on your firewalls, your print servers, and your DNS activity.
    The ever growing list of Lire-supported services clearly outstrips any
    other software, in large part thanks to the numerous volunteers who
    have pioneered many new services and features. Lire is a total solution
    for your log analysis needs.
}
{http://logreport.org/lire.html}
{2009/01/11}
\fussy{}

@@

    \item The pair of correspondents that exchanged the highest number of
        mails.

### 2009-02-23 doc/logparser-results.tex
Remainder of the corrections to the background chapter.
The beginning of the architecture rules section.
Added a graph that utilises the data about the number of rules tried, lines     parsed, etc. Added examples in background section to deal with remaining XXX items. Dealt with some other XXX items, and added some more. Improved the Makefile and fixed some dependency mistakes. Started adding "last checked" dates to all URLs. Improved the introduction in the architecture chapter.

@@

@@

@@

\subsection{Scalability as the Number of Rules Rises}

@@

>>>>

\label{Results}

XXX SHOULD THIS CHAPTER BE CALLED ``Results'' OR ``Evaluation''\@?

@@

XXX EXPAND THIS PARAGRAPH\@; THERE MAY BE CONTENT IN THE PAPER\@.  REFER TO
THE NEW GRAPH BELOW TOO\@.

@@

\showgraph{build/graph-rules-tried-per-log-line}{Mean number of rules tried
per log line}{Mean number of rules tried per log line}

@@

\subsection{Scalability When The Number Of Rules Rises}

@@

\newpage{} % XXX

### 2009-02-23 doc/logparser.tex
Improved the framework section in the architecture chapter; it still needs work, but I'm moving on to actions.
Updated rule numbers.
Enable pdflatex's kerning features for a better looking document.
The beginning of the architecture rules section.
Added a graph that utilises the data about the number of rules tried, lines     parsed, etc. Added examples in background section to deal with remaining XXX items. Dealt with some other XXX items, and added some more. Improved the Makefile and fixed some dependency mistakes. Started adding "last checked" dates to all URLs. Improved the introduction i

% $Id$
% XXX WOULD A SMALLER FONT BE BETTER?
\documentclass[a4paper,12pt,draft]{report}
\pagestyle{headings}

% Useful stuff for math mode.
\usepackage{amstext}
% Include images
\usepackage[final]{graphicx}
\usepackage{url}
% Add the bibliography into the table of contents.
\usepackage[chapter,numbib]{tocbibind}
% Create a label for the last page.  Might be useful for "page 23/79" or
% something.
\usepackage{lastpage}

% Extra footnote functionality, including references to earlier footnotes.
\usepackage[bottom]{footmisc}
% The Euro symbol
\usepackage{eurofont}

@@

% Tell Latex to use scalable fonts
\usepackage{type1cm}

@@

@@

\newcommand{\numberOFrules}[0]{%
    169%
}

\newcommand{\numberOFrulesMINIMUM}[0]{%
    115%
}

% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    68.04\%%
}

% \numberOFrules as percentage increase of \numberOFrulesMINIMUM
\newcommand{\numberOFrulesMAXIMUMpercentage}[0]{%
    46.95\%%
}

@@

@@

XXX WHY DOES \\ref{} NOT WORK PROPERLY FOR CHAPTERS IN THE APPENDIX\@?

@@

xXX IMPROVE BIBLIOGRAPHY APPEARANCE\@: NAMES LOOK UGLY (E.G.\ P.\
\empty{}.\ Porters)\@.

XXX CHECK WITH CARL ABOUT CAPITALISATION IN SECTION AND SUBSECTION
HEADINGS\@.

XXX MAYBE DO SOMETHING ABOUT underfull hboxes IN BIBLIOGRAPHY WHEN URLS
WRAP\@?

>>>>

% $Id$
\documentclass[a4paper,12pt,draft]{report}
\pagestyle{headings}

% Useful stuff for math mode.
\usepackage{amstext}
% Include images
\usepackage[final]{graphicx}
% Add the bibliography into the table of contents.
\usepackage[chapter,numbib]{tocbibind}
% Create a label for the last page.  Might be useful for "page 23/79" or
% something.
\usepackage{lastpage}

% Extra footnote functionality, including references to earlier footnotes.
% Removed for now, because it generates a warning:
% "LaTeX Warning: Command \@makecol  has changed."
%\usepackage[bottom]{footmisc}
% The Euro symbol
\usepackage{eurofont}

@@

% Tell Latex to use scalable fonts
\usepackage{type1cm}
% Enable nice kerning features.
\usepackage[final]{microtype}

@@

% Add last checked dates to all URLs
\newcommand{\urlLastChecked}[2]{%
    \url{#1} (last checked #2)%
}

@@

\newcommand{\numberOFrules}[0]{%
    185%
}

\newcommand{\numberOFrulesMINIMUM}[0]{%
    116%
}

% \numberOFrulesMINIMUM as percentage of \numberOFrules
\newcommand{\numberOFrulesMINIMUMpercentage}[0]{%
    62.70\%%
}

% \numberOFrules as percentage increase of \numberOFrulesMINIMUM
\newcommand{\numberOFrulesMAXIMUMpercentage}[0]{%
    59.48\%%
}

@@

\newcommand{\numberOFactions}[0]{%
    18%
}

\newcommand{\numberOFruleINTERSECTIONS}[0]{%
    14196%
}

@@

XXX WHY DOES $\backslash{}$ref\{\} NOT WORK PROPERLY FOR CHAPTERS IN THE APPENDIX\@?

@@
